\Title

### 2022-09-29 Grimaldi-etal-BiolCybernetics.tex
Introduction: 1.3

Neuroscience could provide revolutionary answers to these two closely related problems. Indeed, the human brain has the remarkable property of being able to react at any time, while consuming a reasonable amount of energy, estimated at 5W at rest. This system is the result of millions of years of natural selection, and a striking difference between biological neural networks and artificial ones is the representation they use. CNNs, for example, represent information passing from one layer to another as tensors representing visual information densely over the visual topography, while representing different properties in different channels. These networks have been known to mimick many propoerties of biological systems~\citep{Schrimpf2020}, yet, this score does not count rapidity or energy. On the other hand, information in a large majority of biological neural networks is represented as spikes, i.e. prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that emitted this spike. Why spikes? It is not yet known why most biological systems have evolved to use this representation, but one major advantage seems to be some form of efficiency. In the human brain, each of the approximately 90 trillion nerve cells emits on average one spike per second which fan out to an average of 10000 synapses. Importantly, each neuron fires with great variability, and up to 300 spikes per second. Many attempts have been made to build artificial spiking neural networks, and in particular to achieve the same performance as observed in classical CNNs with a lower energy budget. However, none of these networks has yet surpassed the state of the art performances. Here, we will try to focus on one core property of spiking neurons inspired by neurobiology to progress in this direction.
%
\subsection{What can neuroscience bring to computer vision?}% and the heterogeneous delays in spiking neurons} ~\citep{}
%%%-----------------------------------------------------------------
Let us first review some existing spiking neural network models. A first motivation for these models was to understand biological observations at the different scales of the central nervous system. For example, the Hodgkin-Huxley model describes the dynamics of the membrane potential of a single nerve cell and specifically explains the role of the different voltage-dependent ion channels. It also explains how the model can produce a spike. Larger scale models introduce an simpler dynamics and an explicit spiking mechanism, for example in the Integrate-and-Fire model. It predicts the emergent states of a recurrently connected neural network~\citep{brunel_phase_2000}. Several models have increased the scale of the system to include more neurons, such as The Virtual Brain (TVB)~\citep{sanz_leon_virtual_2013}, which mixes different scales and includes spiking neuron networks with (analog) neural mass models. However, these methods are mainly descriptive with applications in neuroscience but few spiking neurall models have had an application in computer vision. Another class of approaches have designed normative models of spiking neuron networks that aim to have applications in machine learning. One of these is the SpikeNet algorithm which uses a radical approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}. Another type of spiking neural network using precise spike timing attempts to determine the structure of the network in order to minimize a cost function. This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been extanded in novel gradient-based methods. Surrogate gradient for instance is now widely used in methods that attempt to transfer performance from CNNs to SNNs~\citep{zenke}. However, the performance of SNNs is still lower than that of analog networks and the question of the advantage of using spikes remains open.

@@

This paper explores the possibility that spiking motifs may be a useful representation used in neurobiology and that we may transsfer this to neuromorphic algorithms. Our spiking neural network model is defined as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights. Here, we propose to extend our model to a layer of spiking neurons which include, in addition to synaptic weights, heterogeneous delays. In particular, one afferent may be connected with multiple delays and, crucially, we will explicitly use the delay in the computational process. As a consequence, this extends our previous model by encoding spiking motifs as 3D images. The objective in this model, by including the dimension of temporal delays, is to increase the representational capacities of the classifier. In the perspective of building energy-efficient algorithms, we will also titrate quantitatively the best trade-off between robustness and computation time when increasing the number of these heterogeneous delays. %

@@

In this work, we study the emergence of such spatiotemporal spike motifs when training a single layer of spiking neurons on a supervised classification task (\seeFig{motion_task}). The paper is organized as follows. We develop an SNN-like method capable of learning heterogeneous delays to perform motion detection on an event-based synthetic dataset.   We will first detail the methodology by defining the core spike mechanism using heterogeneous delays. We will then formally define the computational task of motion detection. This will allow us to formalize the final spiking neuron used to learn the model and test its effectiveness.  In the results section, we will detail the emergent structure of the model solving the task and show in particular the similarities with neurobiological observations. We will also test the network with classical stimuli used in neuroscience, and in particular the frequency or contrast responses.  We will also study the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning. This will allow us to show how such a model can provide an effective solution to the energy/accuracy trade-off.  Finally, we will conclude by highlighting the main contributions of this paper, while defining some limitations. This will open up perspectives for future SNNs.  As neuromorphic devices are, by design, good candidates for integrating computations over time, we highlight the fact that this event-driven algorithm is transferable to this type of hardware.

>>>>

Neuroscience could provide revolutionary answers to these two closely related problems. Indeed, the human brain has the remarkable property of being able to react at any time, while consuming a reasonable amount of energy, estimated at 5W at rest. This system is the result of millions of years of natural selection, and a striking difference between biological neural networks and artificial ones is the representation they use. CNNs, for example, represent information passing from one layer to another as tensors representing visual information densely over the visual topography, while representing different properties in different channels. These networks have been known to mimick many propoerties of biological systems~\citep{Schrimpf2020}, yet, this score does not count rapidity or energy. On the other hand, information in a large majority of biological neural networks is represented as spikes, i.e. prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that emitted this spike. Why spikes? It is not yet known why most biological systems have evolved to use this representation, but one major advantage seems to be some form of efficiency. In the human brain, each of the approximately 90 trillion nerve cells emits on average one spike per second which fan out to an average of 10000 synapses. Importantly, each neuron fires with great variability, and up to 300 spikes per second. Many attempts have been made to build artificial spiking neural networks, and in particular to achieve the same performance as observed in classical CNNs with a lower energy budget. However, none of these networks has yet surpassed the state of the art performances. Here, we will try to focus on one core property of spiking neurons inspired by neurobiology to progress in this direction.
%
\subsection{What can neuroscience bring to computer vision?}% and the heterogeneous delays in spiking neurons} ~\citep{}
%%%-----------------------------------------------------------------
Let us first review some existing spiking neural network models. A first motivation for these models was to understand biological observations at the different scales of the central nervous system. For example, the Hodgkin-Huxley model describes the dynamics of the membrane potential of a single nerve cell and specifically explains the role of the different voltage-dependent ion channels. It also explains how the model can produce a spike. Larger scale models introduce an simpler dynamics and an explicit spiking mechanism, for example in the Integrate-and-Fire model. It predicts the emergent states of a recurrently connected neural network~\citep{brunel_phase_2000}. Several models have increased the scale of the system to include more neurons, such as The Virtual Brain (TVB)~\citep{sanz_leon_virtual_2013}, which mixes different scales and includes spiking neuron networks with (analog) neural mass models. However, these methods are mainly descriptive with applications in neuroscience but few spiking neurall models have had an application in computer vision. Another class of approaches have designed normative models of spiking neuron networks that aim to have applications in machine learning. One of these is the SpikeNet algorithm which uses a radical approach by encoding information using one spike per neuron~\citep{delorme_spikenet_1999}. Another type of spiking neural network using precise spike timing attempts to determine the structure of the network in order to minimize a cost function. This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been extanded in novel gradient-based methods. Surrogate gradient for instance is now widely used in methods that attempt to transfer performance from CNNs to SNNs~\citep{zenke_remarkable_2021}. However, the performance of SNNs is still lower than that of analog networks and the question of the advantage of using spikes remains open.

@@

This paper explores the possibility that prototypical spiking motifs may be a useful representation used in neurobiology and that we may transfer this hypothesis to neuromorphic algorithms. Our spiking neural network model is defined as a simple layer of leaky integrate-and-fire neurons whose connections are defined by a matrix of synaptic weights. Here, we propose to extend our model to a layer of spiking neurons which include, in addition to synaptic weights, heterogeneous delays. In particular, one afferent may be connected with multiple delays and, crucially, we will explicitly use the delay in the computational process. As a consequence, this extends our previous model by encoding spiking motifs as 3D images. The objective in this model, by including the dimension of temporal delays, is to increase the representational capacities of the classifier. In the perspective of building energy-efficient algorithms, we will also titrate quantitatively the best trade-off between robustness and computation time when increasing the number of these heterogeneous delays. %

@@

In this work, we study the emergence of such spatiotemporal spike motifs when training a single layer of spiking neurons on a supervised classification task (\seeFig{motion_task}). The paper is organized as follows. We develop a theroretically-defined SNN capable of learning heterogeneous delays which we first evaluate on an event-based synthetic dataset. We will first detail the methodology by defining the core spike mechanism using heterogeneous delays. We will then formally define the computational task of motion detection. This will allow us to formalize the  spiking neuron used to learn the model and test its effectiveness. We aim at applying this model to a real-world task of motion detection which we will first define. Then we will show how the model can be adapted to solve the motion detection task. In the results section, we will detail the emergent structure of the model solving the task and show in particular the similarities with neurobiological observations. We will also test the network with classical stimuli used in neuroscience, and in particular the frequency or contrast responses.  We will also study the efficiency of the motion detection mechanism and in particular its resilience to synaptic weight pruning. This will allow us to show how such a model can provide an effective solution to the energy/accuracy trade-off.  Finally, we will conclude by highlighting the main contributions of this paper, while defining some limitations. This will open up perspectives for future SNNs.  In particular, as neuromorphic devices are by design good candidates for integrating computations over time, we highlight the fact that this event-driven algorithm is transferable to this type of hardware.
